---
title: "Lab3"
author: "Zihao Han"
date: "February 27, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
HAPPY <- readRDS("HAPPY.rds")
head(HAPPY)
View(HAPPY)
```
#Data Cleaning
## Data clean 1: replace missing value "IAP, "DK and "NA to "NA"
```{r}
HAPPY[is.na(HAPPY)] <- NA
HAPPY[HAPPY == "NA"] <- NA
HAPPY[HAPPY == "IAP"] <- NA
HAPPY[HAPPY == "DK"] <- NA
```

## Data clean 2: Check the type of the variable and cast into the right type 
##### After the data clean, the AGE variable shows as numeric.

```{r}
HAPPY$AGE[HAPPY$AGE == "89 OR OLDER"] <- 89
HAPPY$AGE <- as.numeric (HAPPY$AGE)
is.numeric (HAPPY$AGE)
str (HAPPY)
```
## Data clean 3: Bring all levels of factors into a sensible order based on MARITAL and AGE

```{r}
HAPPY <- HAPPY[order(HAPPY$MARITAL, HAPPY$AGE),]
head (HAPPY, 20)
```
##Investigate the distributions of 2 different variables.
```{r}
HAPPY <- na.omit (HAPPY)
summary(HAPPY)
library(ggplot2)
ggplot(HAPPY,aes(x=HAPPY,fill=HAPPY,color=HEALTH))+geom_bar(fill="gray",width=0.95)+facet_wrap(~SEX,ncol=30,scales =  "free_y")+xlab("Happiness")+theme_bw()+theme(text = element_text(size=8))

```
##  Investigate the relationship between happiness and three other variables in the data.
```{r}
library(ggmosaic)
ggplot(data = HAPPY)+geom_mosaic(aes(x=product(HAPPY),fill=MARITAL))
ggplot(HAPPY,aes(x=HAPPY,y=WTSSALL,color= SEX))+geom_point()
ggplot(HAPPY,aes(x=HAPPY,y=as.numeric(AGE),color=SEX))+geom_boxplot()+coord_flip()
ggplot(data = HAPPY)+geom_mosaic(aes(x=product(HAPPY),fill=DEGREE))
```

## Data Eploration 3: What do you learn from the data set? Write one or two paragraphs to describe.

##### by graphing the relation between each vairabel, we could say that people who haver better health will have more happiness, and people who married will be more happy about their life. half of 40-ish people is happy and half is not happy(EXPECIALLY 40-IHS WOMAN IS NOT HAPPY). and peopel who have highschool or higher degree will be more happier.