---
title: "individual1.Rmd"
author: "xiwang"
date: "02/19/19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Open file
```{r}
HAPPY <- readRDS("HAPPY.rds")
dim (HAPPY)
```

#Data Cleaning
## Data clean 1: replace missing value "IAP, "DK" and "NA" to NA

```{r}
HAPPY[is.na(HAPPY)] <- NA
HAPPY[HAPPY == "NA"] <- NA
HAPPY[HAPPY == "IAP"] <- NA
HAPPY[HAPPY == "DK"] <- NA
```

## Data clean 2: Check the type of the variable and cast into the right type 
##### After the data clean, the AGE variable shows as numeric.

```{r}
HAPPY$AGE[HAPPY$AGE == "89 OR OLDER"] <- 89
HAPPY$AGE <- as.numeric (HAPPY$AGE)
is.numeric (HAPPY$AGE)
str (HAPPY)
```

## Data clean 3: Bring all levels of factors into a sensible order based on MARITAL and AGE

```{r}
HAPPY <- HAPPY[order(HAPPY$MARITAL, HAPPY$AGE),]
head (HAPPY, 20)
```

#Data Exploration

## Data Eploration 1: Investigate the distributions of 2 different variables.

```{r}
# Remove missing value in the data frame
HAPPY <- na.omit (HAPPY)
dim (HAPPY)

# Explore the dataset
library (ggplot2)
# Check the distribution of variable: HAPPY based on age
ggplot (HAPPY, aes(x=HAPPY, fill= HAPPY)) + geom_bar() + facet_grid(~SEX, scales="free_y") 

# Check the distribution of variable: AGE
ggplot (HAPPY, aes(x=AGE, color= "red")) + geom_histogram (position = "stack" , binwidth = .5)

# Check the distribution of variable HEALTH: 
ggplot (HAPPY, aes(x=HEALTH, fill=HEALTH)) + geom_bar() + theme_dark(base_size = 11, base_family = "")

```


## Data Eploration 2: Investigate the relationship between happiness and three other variables in the data.

```{r}
library(ggmosaic)
head (HAPPY)
# Relationship between HAPPY to AGE
ggplot (HAPPY, aes(x=HAPPY, y=AGE)) + geom_boxplot(size=1)

# Relationship between HAPPY to MARITAL
ggplot(data = HAPPY)  + 
  geom_mosaic(aes(x = product(HAPPY), fill=MARITAL))  + ylab ("MARITAL") + xlab ("HAPPY") + coord_flip()

# Relationship between HAPPY to MARITAL
ggplot(data = HAPPY)  + 
  geom_mosaic(aes(x = product(HAPPY), fill=HEALTH)) + ylab ("HEALTH") + xlab ("HAPPY") + coord_flip()

```


## Data Eploration 3: What do you learn from the data set? Write one or two paragraphs to describe.

##### 1. First, there are so many missing value in this dataset, after cleaning it, almost half of the data were removed.
##### 2. In terms of the variable HAPPY, there are more people reported they are in Pretty happy and Very happy conditions.
##### 3. There is a right skewed distribution of participant AGE in this dataset.
##### 4. In terms of the variable HEALTH, there are more people reported they are in Excellent and Good conditions.
##### 5. From the boxplot, there is no significant difference can bd identified between the AGE and HAPPY.
##### 6. There are more participants are in a Married status, and those people are have higher percentages in both Very Happy and Pretty Happy conditions.
##### 7. Both Divorced and Widowed particpants showed a large percentages of Not too Happy than other people.
##### 8. Participants in a very Excellent health contion shows a larger percent of Very Happy than other groups, while participants in a Poor health contion shows a larger percentage of Not too Happy than other groups.





