---
title: "Lab3"
author: "Zihao Han"
date: "February 27, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

HAPPY <- readRDS("HAPPY.rds")
head(HAPPY)
View(HAPPY)
```
#Data Cleaning
## Data clean 1: replace missing value "IAP, "DK and "NA to "NA"
```{r}
HAPPY[is.na(HAPPY)] <- NA
HAPPY[HAPPY == "NA"] <- NA
HAPPY[HAPPY == "IAP"] <- NA
HAPPY[HAPPY == "DK"] <- NA
```

## Data clean 2: Check the type of the variable and cast into the right type 
##### After the data clean, the AGE variable shows as numeric.

```{r}
HAPPY$AGE[HAPPY$AGE == "89 OR OLDER"] <- 89
HAPPY$AGE <- as.numeric (HAPPY$AGE)
is.numeric (HAPPY$AGE)
str (HAPPY)
```
## Data clean 3: Bring all levels of factors into a sensible order based on MARITAL and AGE

```{r}
HAPPY <- HAPPY[order(HAPPY$MARITAL, HAPPY$AGE),]
head (HAPPY, 20)
```
##Investigate the distributions of 2 different variables.
```{r}
HAPPY <- na.omit (HAPPY)
summary(HAPPY)
library(ggplot2)
ggplot(HAPPY,aes(x=HAPPY,fill=HAPPY,color=HEALTH))+geom_bar(fill="gray",width=0.95)+facet_wrap(~SEX,ncol=30,scales =  "free_y")+xlab("Happiness")+theme_bw()+theme(text = element_text(size=8))

```
##  Investigate the relationship between happiness and three other variables in the data.
```{r}
library(ggmosaic)
ggplot(data = HAPPY)+geom_mosaic(aes(x=product(HAPPY),fill=MARITAL))
ggplot(HAPPY,aes(x=HAPPY,y=WTSSALL))+geom_point()
ggplot(HAPPY,aes(x=HAPPY,y=as.numeric(AGE)))+geom_boxplot()+coord_flip()
ggplot(data = HAPPY)+geom_mosaic(aes(x=product(HAPPY),fill=DEGREE))
```

## Data Eploration 3: What do you learn from the data set? Write one or two paragraphs to describe.

##### 1. First, there are so many missing value in this dataset, after cleaning it, almost half of the data were removed.
##### 2. In terms of the variable HAPPY, there are more people reported they are in Pretty happy and Very happy conditions.
##### 3. There is a right skewed distribution of participant AGE in this dataset.
##### 4. In terms of the variable HEALTH, there are more people reported they are in Excellent and Good conditions.
##### 5. From the boxplot, there is no significant difference can bd identified between the AGE and HAPPY.
##### 6. There are more participants are in a Married status, and those people are have higher percentages in both Very Happy and Pretty Happy conditions.
##### 7. Both Divorced and Widowed particpants showed a large percentages of Not too Happy than other people.
##### 8. Participants in a very Excellent health contion shows a larger percent of Very Happy than other groups, while participants in a Poor health contion shows a larger percentage of Not too Happy than other groups.

